Automated data-driven symbolic modeling of the lift forces acting on flapping wings

### 2021-02-07 paper/main.tex
Minor edit to first paragraph in Results section
Add TODO
Edit data-driven models subsection

% TODO: Reword this section to not focus on the number 58 right off the bat.
\subsection*{Data-driven models}
The strategy for modeling lift forces symbolically was to train candidate
equations on 58 of the 68 total experiments simultaneously, with the remaining
ten reserved for testing.  These 58 training sets were intended to represent a
diverse sampling of wing geometry, flapping speed and amplitude combinations.
The results from these 58 experiments were compiled into a single dataset
listing roughly two periods of smoothed, registered flapping data per
experiment.  The data consisted of instantaneous measured vertical force along
with the measured value of each kinematic variable, the first and second
derivatives of those variables, and the geometric parameters necessary to
describe the wing's shape and distribution of area, totaling 31 separate
variables. By training on 58 experiments simultaneously, the symbolic regression
algorithm generated equations that fit all of the data, using the some, but not
all, of the provided variables to account for variations in kinematics and wing
geometry between experiments. The search was conducted for approximately 24
hours, using 60 cores on a computing cluster.

Table~\ref{table:eqmodels} lists the performance of five of the leading
candidate models, along with five analytical models for comparison. We measure
performance as the mean absolute error (MAE) between predicted and measured
vertical force in the 10 test set experiments withheld from training. Our
primary results, provided in the column labeled ``MAE Not Fitted'', use
coefficient values for all models that were optimized on the training set.
However, since the accuracy of the analytical models may depend on coefficients
that vary with wing shape, flapping speed, or other factors, we also provide
secondary results in the column labeled ``MAE Fitted'', which use coefficient
values that were fitted to the test set experiments.

% TODO: Make sure to define all variables used in these tables.
% TODO: Figure out why we're calling it "U-squared".
% TODO: Decide on better terms than "fitted" and "not fitted".
\begin{table}[ht]
\small
\centering
\begin{tabular}{|l|l|r|r|r|}
\hline
  {\bf Model} & {\bf Terms/Formula} & {\bf Size} & \begin{tabular}{@{}c@{}}{\bf MAE (N)} \\
    Fitted\end{tabular} & \begin{tabular}{@{}c@{}}{\bf MAE (N)} \\ Not Fitted\end{tabular}\\
\hline
\(EQ _5\) & \(C_1RY_{CM}\dot{\psi}^2\ + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 12 & 8.79\e{-4} & 1.19\e{-3}\\
\(EQ _4\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 14 & 8.66\e{-4} & 1.18\e{-3}\\
\(EQ _3\) & \(C_1RY_{CM}\dot{\psi}^2\cos(\psi-\psi^2) + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 16 & 8.49\e{-4} & 1.14\e{-3}\\
\(EQ _2\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\dot{\phi}\dot{\psi}\psi - C_4R\ddot{\theta}\) & 19 & 8.15\e{-4} & 1.14\e{-3}\\
\(EQ _1\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\dot{\phi}\psi - C_4R\dot{\phi}\dot{\psi}\psi - C_5R\ddot{\theta}\) & 24 & 7.73\e{-4} & 1.12\e{-3}\\
\hline
\(U^2\) Planar 1 & \(L_{planar},  F_{inertial\:(pendulum\:only)}\) & 32 & 1.38\e{-3} & 1.47\e{-3}\\
\(U^2\) Planar 2 & \(L_{planar}, F_{inertial}\) & 46 & 1.18\e{-3} & 1.29\e{-3}\\
\(U^2\) Total & \(L_{trans}, F_{inertial}\) & 96 & 1.18\e{-3} & 1.29\e{-3}\\
WW~\cite{whitney2010aeromechanics} & \(L_{trans}, AM_{trans}, AM_{rot}\) & 121 & 1.02\e{-3} & 1.23\e{-3}\\
PW~\cite{pesavento2004falling} & \(L_{trans}, L_{rot}, AM_{trans}, AM_{trans-rot}\) & 130 & 1.01\e{-3} & 1.20\e{-3}\\
\hline
\end{tabular}
\caption{\label{table:eqmodels}Comparison of data-driven models and analytical
  equations of lift.  The models are sorted by equation size, which is
  calculated as the sum of arithmetic and trigonometric operators required to
  compute a force prediction from kinematic and geometric data.  MAE reports
  average mean absolute error (in Newtons of vertical force) for the ten test
  experiments.}
\end{table}

>>>>

% TODO: Reword this section to not focus on the number 58 right off the bat.
\subsection*{Data-driven models}
We trained a set of data-driven models to predict the instantaneous vertical
force produced by flapping wings as a function of the kinematic variables and
derivatives describing their motion, and geometric parameters describing the
wing shape, totaling 31 separate variables. We trained these models on data from
58 experiments, selected at random from our 68 total experiments, which
represent a diverse sampling of wing geometry, flapping speed and amplitude
combinations. The remaining 10 experiments were reserved for testing. Models
were trained using the symbolic regression algorithm developed by Schmidt and
Lipson \cite{schmidt2009distilling} to minimized the mean absolute error (MAE)
between measured and predicted vertical force. The training process ran for
approximately 24 hours using 60 cores on a computing cluster.

By training on data representing a range of wing shapes and flapping speeds and
amplitudes, the model search effectively narrowed the large set of input
variables to the most preditive information and only utilized those terms that
accounted for variations in kinematics and wing geometry between experiments.
Table~\ref{table:eqmodels} lists the performance of five of the leading
candidate models, along with five analytical models for comparison, as measured
by MAE in the 10 test set experiments withheld from training. These five models
were selected from the Pareto frontier between MAE and formula complexity when
training was terminated. The data-driven models presented here are functions of
the kinematic variables ($\phi$, $\psi$, and $\theta$) and their derivatives,
the wing span ($R$), the span-wise and chord-wise locations of the center of
wing area ($X_{CM}$ and $Y_{CM}$), and constant coefficients. The analytical
models for comparison will be described in the following section.

Our primary results, provided in the column labeled ``MAE Not Fitted'', use
fixed coefficient values for all models that were jointly optimized on the
training set.  However, since the accuracy of the analytical models may depend
on coefficients that vary with wing shape, flapping speed, or other factors, we
also provide secondary results in the column labeled ``MAE Fitted'', which use
coefficient values that were fitted individually to each of the test set
experiments. In both categories of results, the data-driven models are more
accurate, demonstrating that the relative performance of data-driven models is
not simply due to the choice or optimization of coefficient values.

% TODO: Make sure to define all variables used in these tables.
% TODO: Figure out why we're calling it "U-squared".
% TODO: If U-squared is "velocity squared" then  make sure that's consistent
% with the use of $v^2$ in the L_{trans} formula (equation 1) and elsewhere.
% TODO: Decide on better terms than "fitted" and "not fitted".
% TODO: Explain that the 5 EQ models were taken from the Pareto frontier of
% complexity vs. accuracy at the time of convergence.
% TODO: Explain that C_1,...,C_5 are not the same in each EQ. Maybe do C_i,_j
\begin{table}[ht]
\small
\centering
\begin{tabular}{|l|l|r|r|r|}
\hline
  {\bf Model} & {\bf Terms/Formula} & {\bf Size} & \begin{tabular}{@{}c@{}}{\bf MAE (N)} \\
    Fitted\end{tabular} & \begin{tabular}{@{}c@{}}{\bf MAE (N)} \\ Not Fitted\end{tabular}\\
\hline
\(EQ _5\) & \(C_1RY_{CM}\dot{\psi}^2\ + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 12 & 8.79\e{-4} & 1.19\e{-3}\\
\(EQ _4\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 14 & 8.66\e{-4} & 1.18\e{-3}\\
\(EQ _3\) & \(C_1RY_{CM}\dot{\psi}^2\cos(\psi-\psi^2) + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 16 & 8.49\e{-4} & 1.14\e{-3}\\
\(EQ _2\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\dot{\phi}\dot{\psi}\psi - C_4R\ddot{\theta}\) & 19 & 8.15\e{-4} & 1.14\e{-3}\\
\(EQ _1\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\dot{\phi}\psi - C_4R\dot{\phi}\dot{\psi}\psi - C_5R\ddot{\theta}\) & 24 & 7.73\e{-4} & 1.12\e{-3}\\
\hline
\(U^2\) Planar 1 & \(L_{planar},  F_{inertial\:(pendulum\:only)}\) & 32 & 1.38\e{-3} & 1.47\e{-3}\\
\(U^2\) Planar 2 & \(L_{planar}, F_{inertial}\) & 46 & 1.18\e{-3} & 1.29\e{-3}\\
\(U^2\) Total & \(L_{trans}, F_{inertial}\) & 96 & 1.18\e{-3} & 1.29\e{-3}\\
WW~\cite{whitney2010aeromechanics} & \(L_{trans}, AM_{trans}, AM_{rot}\) & 121 & 1.02\e{-3} & 1.23\e{-3}\\
PW~\cite{pesavento2004falling} & \(L_{trans}, L_{rot}, AM_{trans}, AM_{trans-rot}\) & 130 & 1.01\e{-3} & 1.20\e{-3}\\
\hline
\end{tabular}
\caption{\label{table:eqmodels}Comparison of data-driven models and analytical
  equations of lift.  The models are sorted by equation size, which is
  calculated as the sum of arithmetic and trigonometric operators required to
  compute a force prediction from kinematic and geometric data.  MAE reports
  average mean absolute error (in Newtons of vertical force) for the ten test
  experiments.}
\end{table}
